\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-07-17 syntax.tex
Add overview of syntax pieces.

@@

$$
\Rule[name=Bool]
{}
{\goodTerm{\Gamma}{"tt", "ff"}{\cB}}
\quad
\Rule[]
{\goodSub{\Gamma}{\gamma}{\Delta}}
{\goodType{\Gamma}{(El \ T)[\gamma] \equiv El \ (T[\gamma]) }{j} \quad
 \goodTerm{\Gamma}{"tt"[\gamma] \equiv "tt"}{\cB} \quad 
 \goodTerm{\Gamma}{"ff"[\gamma] \equiv "ff"}{\cB} 
}
$$
\judgebox{\goodSub{\Gamma}{\sigma}{\Delta}}
$$
\Rule[name=Ept Subst]
{}{\goodSub{\Gamma}{\epsilon}{\cdot}}
\quad
\Rule[]
{}{\goodSub{\Gamma}{"id"}{\Gamma}}
\quad
\Rule[name=Sub Comp]{
  \goodSub{\Delta}{\delta}{\Theta}
  \quad \goodSub{\Gamma}{\gamma}{\Delta} 
}{\goodSub{\Gamma}{\delta \circ \gamma}{\Theta}}
\Rule[name=Sub Extend]
{\goodSub{\Gamma}{\sigma}{\Delta} \quad \goodTerm{\Gamma}{t}{A[\sigma]}}
{\goodSub{\Gamma}{(\sigma, t)}{(\Delta, A)}}
$$

$$
\Rule[name=Proj Subst-1]
{\goodSub{\Gamma}{\sigma}{(\Delta, A)}}
{\goodSub{\Gamma}{\pi_1 \sigma}{\Delta}}
\quad
\Rule[name=Proj Subst-2]
{\goodSub{\Gamma}{\sigma}{(\Delta, A)}}
{\goodSub{\Gamma}{\pi_2 \sigma}{A[\pi_1 \sigma]}}
\quad
\Rule[name=Proj-Ext]
{}
{\goodSub{\Gamma}{(\pi_1 \sigma, \pi_2 \sigma) \equiv \sigma}{\Delta}}
$$

@@

For these two types using binder, de Bruijn indices emphasizes its difference from the conventional setting---we directly use $\lambda t$ as a function term where $t$ is the body, no binder indication is required---to refer to variables in the context, we use $\pi_1,\pi_2$ from substitution calculus instead of using their names in the conventional setting.  For example, the identity function $\lambda x. x$ is actually $\lambda(\pi_2 id)$. $"app"~t$ is even more different than that of the original setting---we don't require an application argument now. The reason is, (1) we asked for argument in conventional setting so we can use meta-level substitution as part of the dynamic, but here we use explicit substitution. If the user want to substitute some arguments, they can just use $[\gamma]$ instead. (2) Formulated in this way, "app" and $\lambda$ is symmetrical inverse to each other (just like how $El$ and "c" are to each other). This makes the $\beta\eta$-rule elegantly straightforward. Besides $\lambda$, $\Pi$ and $\Sigma$ are also using binders and are formulated in a similar style.

@@

\subsubsection{\textbf{Core Syntax of the Extended MLTT}}
\begin{figure}[!htb]
  \begin{align*}
&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\tau~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\tau}{n}} \\ &h~:~"Inh"~\Gamma~\sigma_1~\sigma_2&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~T&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}\YZ{A little unfriendly to the reader that there are three different things that sigma can range over. Besides, later you also use gamma to range over substitutions and s to range over W-types, which add to the confusion.}\EDJreply{Fair point. I have fixed fix sigma and tau to be about signatures and W signature, and use gamma as substitutions}
  \end{align*}
\caption{Extended Judgement forms, as Agda types}
\end{figure}
\begin{figure}[!htb]
  \begin{minipage}[b]{0.3\linewidth}
      $$
      \Rule[name=Sig]
      {\goodCtx{\Gamma}{i} \quad n \in \nat}
      {\goodSig{\Gamma}{\_}{n} \text{ is well-formed}}
      $$
      $$
      \Rule[name=Seal]
      {\goodCtx{\Gamma}{i} \quad \goodSig{\Gamma}{\sigma}{n} 
      \quad \goodType{\Gamma}{A}{} }
      {\goodSeal{\Gamma}{\_}{\sigma}{A} \text{ is well-formed}}
      $$\YZ{Doesn't the left column show well-formedness rules for ONLY TWO of the four new judgments? Besides, are these rules used anywhere in the following? Seems a bit unusual to have well-formedness rules for judgments per se.}\EDJreply{Yes. Right column show all the well-formedness rule for all the judgement. That is unusual because it is a direct translation from QIIT. }
  \end{minipage}
  \begin{minipage}[b]{0.6\linewidth}
    \begin{minted}[]{agda}
      data Sig  : Con → ℕ → Set
      data WSig : Con → ℕ → Set 
      data Inh  : (Γ : Con) → Sig Γ n → Sig Γ m → Set
      Seal : (Γ : Con) → Sig Γ n → Sig Γ n → Set
    \end{minted}
  \end{minipage}

@@

The core language builds on MLTT and adds four new judgement forms.
% Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is just a
% shorthand for term judgement
% $\goodTerm{\Gamma , \cC \sigma}{\_}{A[\pi_1]}$ where $A$ is a type.
% We have shown some exemplar rules for all four standard types of
% judgments. Now we focus on the newly introduced facility.

We have two kinds of signatures, one for inductive types, the other for linkages. We start with inductive type signature.

>>>>

\begin{figure}[!htb]
  \begin{alignat*}{3}
    \text{Context} \quad & \Gamma, \Delta &&= \cdot~|~ \Gamma, A \\ 
    \text{Types}   \quad & A, B, T        &&= \cU~|~\cB~|~\bot~|~\top~|~\Pi A B~|~\Sigma A B~|~El~t~|~T[\gamma]\\
    \text{Terms}   \quad & t, t_1, t_2    &&= "c"~T~|~t[\gamma]~|~\lambda t~|~"app"~t~|~(t_1,t_2)~|~"pjl"~t~|~"pjr"~t~|~"tt"~|~"ff"~|~\pi_2 \gamma  \\
    \text{Substitution}   \quad & \gamma, \gamma_1, \gamma_2    &&= \epsilon~|~"id"~|~\gamma_1 \circ \gamma_2~|~(\gamma, t)~|~\pi_1 \gamma \\
  \end{alignat*}
\caption[Caption for LOF]{Overview of Instrinsically-typed Syntax for vanilla MLTT\protect\footnotemark}
\end{figure}
\footnotetext{This is not the definition of our syntax, just a grand overview of the intrinsically-typed syntax, to expose the reader with an idea on, for example, whether a syntax piece is a term or type, the general impression of the expressiveness of the calculus}

@@

$$
\Rule[name=Bool]
{}
{\goodTerm{\Gamma}{"tt", "ff"}{\cB}}
\quad
\Rule[]
{\goodSub{\Gamma}{\gamma}{\Delta}}
{\goodType{\Gamma}{(El \ T)[\gamma] \equiv El \ (T[\gamma]) }{j} \quad
 \goodTerm{\Gamma}{"tt"[\gamma] \equiv "tt"}{\cB} \quad 
 \goodTerm{\Gamma}{"ff"[\gamma] \equiv "ff"}{\cB} 
}
$$
\judgebox{\goodSub{\Gamma}{\sigma}{\Delta}}
$$
\Rule[name=Ept Subst]
{}{\goodSub{\Gamma}{\epsilon}{\cdot}}
\quad
\Rule[]
{}{\goodSub{\Gamma}{"id"}{\Gamma}}
\quad
\Rule[name=Sub Comp]{
  \goodSub{\Delta}{\delta}{\Theta}
  \quad \goodSub{\Gamma}{\gamma}{\Delta} 
}{\goodSub{\Gamma}{\delta \circ \gamma}{\Theta}}
\Rule[name=Sub Extend]
{\goodSub{\Gamma}{\gamma}{\Delta} \quad \goodTerm{\Gamma}{t}{A[\sigma]}}
{\goodSub{\Gamma}{(\gamma, t)}{(\Delta, A)}}
$$

$$
\Rule[name=Proj Subst-1]
{\goodSub{\Gamma}{\gamma}{(\Delta, A)}}
{\goodSub{\Gamma}{\pi_1 \gamma}{\Delta}}
\quad
\Rule[name=Proj Subst-2]
{\goodSub{\Gamma}{\gamma}{(\Delta, A)}}
{\goodSub{\Gamma}{\pi_2 \gamma}{A[\pi_1 \gamma]}}
\quad
\Rule[name=Proj-Ext]
{}
{\goodSub{\Gamma}{(\pi_1 \gamma, \pi_2 \gamma) \equiv \gamma}{\Delta}}
$$

@@

For these two types using binder, de Bruijn indices emphasizes its difference from the conventional setting---we directly use $\lambda t$ as a function term where $t$ is the body, no binder indication is required---to refer to variables in the context, we use $\pi_1,\pi_2$ from substitution calculus instead of using their names in the conventional setting.  For example, the identity function $\lambda x. x$ is actually $\lambda(\pi_2 "id")$. $"app"~t$ is even more different than that of the original setting---we don't require an application argument now. The reason is, (1) we asked for argument in conventional setting so we can use meta-level substitution as part of the dynamic, but here we use explicit substitution. If the user want to substitute some arguments, they can just use $[\gamma]$ instead. (2) Formulated in this way, "app" and $\lambda$ is symmetrical inverse to each other (just like how $El$ and "c" are to each other). This makes the $\beta\eta$-rule elegantly straightforward. Besides $\lambda$, $\Pi$ and $\Sigma$ are also using binders and are formulated in a similar style.

@@

\subsubsection{\textbf{Core Syntax of the Extended MLTT}}

\begin{figure}[!htb]
  \begin{alignat*}{3}
    \text{Inductive Type Signature}   \quad & \tau       &&= w\cdot~|~w^+~\tau~A~B~|\tau[\gamma] \\
    \text{Linkage Signature}          \quad & \sigma       &&= \nu\cdot~|~\nu^+~\sigma~\{f\}~T~|~\sigma[\gamma]~|~p_1\nu~\sigma~| \\
    \text{Seal}          \quad & f       &&= t~|~p_f\nu'~\sigma\\    
    \text{Types}   \quad & A, B, T        &&= \bW \tau~|~\pi_1^j~|~\pi_2^j~|~\cL\sigma~|\cC\sigma~|~p_1\nu'~\sigma~|~p_2\nu~\sigma\\
    \text{Terms}   \quad & t, t_1, t_2    &&= "W"~T~|~"Wsup"~T~t_1~t_2~|~\mu\cdot~|~\mu^+~t~\{f\}~T~|~p_1\mu~t~|~p_2\mu~t~|~\cCt~t  \\
  \end{alignat*}
\caption{Overview of Instrinsically-typed Syntax for extended MLTT}
\end{figure}
\begin{figure}[!htb]
  \begin{align*}
&\sigma~:~"Sig"~\Gamma~n&&\mathboxtext{\goodSig{\Gamma}{\sigma}{n}}  &&\tau~:~"WSig"~\Gamma~n&&\mathboxtext{\goodWSig{\Gamma}{\tau}{n}} \\ &h~:~"Inh"~\Gamma~\sigma_1~\sigma_2&&\mathboxtext{\goodInh{\Gamma}{h}{\sigma}{\tau}} && f~:~"Seal"~\Gamma~\sigma~T&&\mathboxtext{\goodSeal{\Gamma}{f}{\sigma}{A}}\YZ{A little unfriendly to the reader that there are three different things that sigma can range over. Besides, later you also use gamma to range over substitutions and s to range over W-types, which add to the confusion.}\EDJreply{Fair point. I have fixed fix sigma and tau to be about signatures and W signature, and use gamma as substitutions}
  \end{align*}
\caption{Extended Judgement forms, as Agda types}
\end{figure}
\begin{figure}[!htb]
  \begin{minipage}[b]{0.3\linewidth}
      $$
      \Rule[name=Sig]
      {\goodCtx{\Gamma}{i} \quad n \in \nat}
      {\goodSig{\Gamma}{\_}{n} \text{ is well-formed}}
      $$
      $$
      \Rule[name=Seal]
      {\goodCtx{\Gamma}{i} \quad \goodSig{\Gamma}{\sigma}{n} 
      \quad \goodType{\Gamma}{A}{} }
      {\goodSeal{\Gamma}{\_}{\sigma}{A} \text{ is well-formed}}
      $$\YZ{Doesn't the left column show well-formedness rules for ONLY TWO of the four new judgments? Besides, are these rules used anywhere in the following? Seems a bit unusual to have well-formedness rules for judgments per se.}\EDJreply{Yes. Right column show all the well-formedness rule for all the judgement. That is unusual because it is a direct translation from QIIT. }
  \end{minipage}
  \begin{minipage}[b]{0.6\linewidth}
    \begin{minted}[]{agda}
      data Sig  : Con → ℕ → Set
      data WSig : Con → ℕ → Set 
      data Inh  : (Γ : Con) → Sig Γ n → Sig Γ m → Set
      Seal : (Γ : Con) → Sig Γ n → Sig Γ n → Set
    \end{minted}
  \end{minipage}

@@

The core language builds on MLTT and adds four new judgement forms.
% Among them, sealing judgement $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is just a
% shorthand for term judgement
% $\goodTerm{\Gamma , \cC \sigma}{\_}{A[\pi_1]}$ where $A$ is a type.
% We have shown some exemplar rules for all four standard types of
% judgments. Now we focus on the newly introduced facility. 
We have two kinds of signatures, one for inductive types, the other for linkages. We start with inductive type signature.
